Meta-analysis

### 2020-02-28 04_MetaAnalysis.Rmd

```{r MEsummary_Libraries2, message=FALSE, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, bomrang, lme4, RColorBrewer)

@@

# Data
slimmer_PM_dat <- read.csv("data/slimmer_PM_clusterdat.csv"
)
```

@@

```{r variance_histograms}
hist(slimmer_PM_dat$yield_error) # yield error is not normally distributed
hist(log(slimmer_PM_dat$yield_error)) # a log transformation shows it is normally distributed 
```

@@

#### Imputing sample variances from Mean squares
```{r imputing_MSE_dataSpread}
slimmer_PM_dat <- read.csv("data/slimmer_PM_clusterdat.csv")
hist(unique(slimmer_PM_dat$Y_Msquare))
hist(log(unique(slimmer_PM_dat$Y_Msquare))) # log mean square has a more normal distribution

@@

## Meta-analysis  
### metafor package  
Let's load the metafor package we are using to analyse the data, then rearrange the factors we want to examine by placing the control treatment first. This way all treatments will be compared to the no spray controls.  
Next we are log transforming the grain yield and calculating the variance from the trial mean squares.  
Finally we will assign factor classes to the main variables in the meta-analysis. Note that variable trial is a combination of:  
   - Trial identifier  
   - Trial year  
   - Trial location  
   - Host genotype  
   - Trial row spacing

@@

summary(lme_PM)
```
&spun;  
&spun;  
To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.
```{r metafor_contrasts}
anova(PM_mv_AI, L = rbind(
   c(0, 1, -1, 0, 0),
   # early vs Recomended
   c(0, 1, 0, -1, 0),
   # early vs recommended plus
   c(0, 1, 0, 0, -1),
   # early vs Late plus
   c(0, 0, -1, 1, 0),
   # Recommended vs recommended_plus
   c(0, 0, -1, 0, 1),
   # Recommended vs Late_Plus
   c(0, 0, 0, -1, 1)
)) # recommended_plus vs Late_plus
```
Results show with the exclusion of the no spray control, none of the treatments are significantly different, however, early applications treatments on average produced lower yields. Let's view this on a plot.

@@

### netmeta package  
Let's analyse the data again using a different statistical approach to see if our outcome with the `metafor` package was robust. The `netmeta` package uses a frequentness approach to the analysis, and focuses on the pairwise comparisons between treatments.

```{r netmeta-analysis}
library(netmeta)

datPM3 <- slimmer_PM_dat %>%
   group_by(trial, spray_management, n) %>%
   summarize(yi_mean = mean(yi),
             vi_mean = mean(vi)) %>%
   ungroup()

>>>>

```{r MEsummary_Libraries2, message=FALSE, include=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, bomrang, lme4, RColorBrewer, netmeta)

@@

# Data
slimmer_PM_dat <- read.csv("cache/slimmer_PM_clusterdat.csv"
)
```

@@

```{r variance_histograms}
hist_usq(slimmer_PM_dat$yield_error) # yield error is not normally distributed
hist_usq(log(slimmer_PM_dat$yield_error)) # a log transformation shows it is normally distributed 
```

@@

#### Imputing sample variances from Mean squares
```{r imputing_MSE_dataSpread}
slimmer_PM_dat <- read.csv("cache/slimmer_PM_clusterdat.csv")
hist_usq(unique(slimmer_PM_dat$Y_Msquare))
hist_usq(log(unique(slimmer_PM_dat$Y_Msquare))) # log mean square has a more normal distribution

@@

## Meta-analysis  
### metafor package  
Let's load the metafor package we are using to analyse the data, then rearrange the factors we want to examine by placing the control treatment first. This way all treatments will be compared to the no spray controls.  
Next we are log transforming the grain yield and calculating the variance from the trial mean squares.  
Finally we will assign factor classes to the main variables in the meta-analysis. Note that variable trial is a combination of:

   * Trial identifier

   * Trial year

   * Trial location

   * Host genotype

   * Trial row spacing

@@

summary(lme_PM)
```

*****s

To make it easier to compare each of the treatments we can compute the meta-analysis contrasts.
```{r metafor_contrasts}
anova(PM_mv_AI, L = rbind(
   c(0, 1, -1, 0, 0),
   # early vs Recomended
   c(0, 1, 0, -1, 0),
   # early vs recommended plus
   c(0, 1, 0, 0, -1),
   # early vs Late plus
   c(0, 0, -1, 1, 0),
   # Recommended vs recommended_plus
   c(0, 0, -1, 0, 1),
   # Recommended vs Late_Plus
   c(0, 0, 0, -1, 1)
)) # recommended_plus vs Late_plus
```
Results show with the exclusion of the no spray control, none of the treatments are significantly different, however, early applications treatments on average produced lower yields. Let's view this on a plot.

@@

### netmeta package  
Let's analyse the data again using a different statistical approach to see if our outcome with the `metafor` package was robust. The `netmeta` package uses a frequentist approach to the analysis and focuses on the pairwise comparisons between treatments.

```{r netmeta-analysis}
datPM3 <- slimmer_PM_dat %>%
   group_by(trial, spray_management, n) %>%
   summarize(yi_mean = mean(yi),
             vi_mean = mean(vi)) %>%
   ungroup()
